Shallow EDSLs and Object-Oriented Programming

### 2016-06-28 paper/sections/Transformations.tex
Code

\paragraph{Desugaring in Scala}
A corresponding Scala implementation is given below:
\lstinputlisting[linerange=4-17]{../src/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
The method \lstinline{desugar} returns a
\lstinline{Circuit} reflecting the nature of transformations.
Implementing it for each case is
straightforward: \emph{fan n} returns itself; $beside\ c_1\ c_2$ returns a new
$beside$ with $c_1$ and $c_2$ desugared; \emph{id n} just mimics the
formula. Circuits are constructed using the helper methods:
\lstinputlisting[linerange=3-5]{../src/desugar/package.scala}%APPLY:COMPANION

%\weixin{define a show function in the and show the result after
%desugaring?}
\paragraph{Desugaring in Haskell}
Interestingly we can port the implementation back to Haskell:
\lstinputlisting[linerange=6-13]{./code/Desugar.hs}%APPLY:DESUGAR
\bruno{Code is cut-off in my pdf! Are the lines correct; beside is not
  fully defined?}
\noindent This code follows well-known encodings of objects into 
\emph{functional} calculi with recursive records~\cite{bruce97comparing}. Thanks 
to Haskell's laziness, the port of recursive definitions such as
\lstinline{beside} is straightforward.

\subsection{Sophisticated Optimizations}
A final example is a more sophisticated type of optimization, 
which requires inspecting multiple internal
representations of abstract values at the same time.
Such optimizations impose extra difficulty to shallow
embeddings.
%, but both
%Cook~\cite{cook09abstraction} and Reynolds~\cite{reynolds94proceduralabstraction} have shown how it 
%is possible to overcome such difficulties. 
%which imposes difficulty to shallow embeddings.
For instance, consecutive \emph{id}s put side by side can be merged together:
$$
beside\ (id\ m)\ (id\ n) = id\ (m + n)
$$
This rule applies to a $beside\ c_1\ c_2$ only when both $c_1$ and $c_2$ are
$id$, also their inner values are extracted for constructing a new $id$.
Fortunately, Scala has built-in support for pattern matching by decorating
classes with \lstinline{case} modifiers or defining extractor objects~\cite{emir2007matching}.
%For extensibility reasons, case classes can not be used
%because case class to case class extension is not supported.
Here, we define the extractor for \lstinline{Id}:
\lstinputlisting[linerange=18-21]{../src/optimizations/Circuit.scala}%APPLY:EXTRACTOR
Then we can implement this optimization:
\lstinputlisting[linerange=4-15]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA

>>>>

\paragraph{Desugaring in Scala}
A corresponding Scala implementation is given below:
\lstinputlisting[linerange=4-17]{../src/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
The method \lstinline{desugar} returns a
\lstinline{Circuit} reflecting the nature of transformations.
Implementing it for each case is
straightforward: \emph{fan n} returns itself; $beside\ c_1\ c_2$ returns a new
$beside$ with $c_1$ and $c_2$ desugared; \emph{id n} just mimics the
formula. Circuits are constructed using the helper methods:
\lstinputlisting[linerange=3-6]{../src/desugar/package.scala}%APPLY:COMPANION

%\weixin{define a show function in the and show the result after
%desugaring?}
\paragraph{Desugaring in Haskell}
Interestingly we can port the implementation back to Haskell:
\lstinputlisting[linerange=6-15]{./code/Desugar.hs}%APPLY:DESUGAR
\noindent This code follows well-known encodings of objects into 
\emph{functional} calculi with recursive records~\cite{bruce97comparing}. Thanks 
to Haskell's laziness, the port of recursive definitions such as
\lstinline{beside} is straightforward.

\subsection{Sophisticated Optimizations}
A final example is a more sophisticated type of optimization, 
which requires inspecting multiple internal
representations of abstract values at the same time.
Such optimizations impose extra difficulty to shallow
embeddings.
%, but both
%Cook~\cite{cook09abstraction} and Reynolds~\cite{reynolds94proceduralabstraction} have shown how it 
%is possible to overcome such difficulties. 
%which imposes difficulty to shallow embeddings.
For instance, consecutive \emph{id}s put side by side can be merged together:
$$
beside\ (id\ m)\ (id\ n) = id\ (m + n)
$$
This rule applies to a $beside\ c_1\ c_2$ only when both $c_1$ and $c_2$ are
$id$, also their inner values are extracted for constructing a new $id$.
Fortunately, Scala has built-in support for pattern matching by decorating
classes with \lstinline{case} modifiers or defining extractor objects~\cite{emir2007matching}.
%For extensibility reasons, case classes can not be used
%because case class to case class extension is not supported.
Here, we define the extractor for \lstinline{Id}:
\lstinputlisting[linerange=18-18]{../src/optimizations/Circuit.scala}%APPLY:EXTRACTOR
Then we can implement this optimization:
\lstinputlisting[linerange=4-15]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA
